<<---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r}

setwd("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 3")

library(plyr)
library(caret)
library(lme4)
library(pROC)

```

## Assignment 3 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia, focusing on pitch.
In the course of this assignment we will use them to try to automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.

### Question 1: Can you diagnose schizophrenia from pitch range only? If so, how well?

Build a logistic regression to see whether you can diagnose schizophrenia from pitch range only.

Calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve) on a logistic regression using the full dataset. Don't forget the random effects!

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures.

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?

```{r}

#Loading data
data = read.csv("final_rqa.csv")

#Removing first column
data = data[, -1]

#Renaming control and schizophrenia to 0 and 1 
data$diagnosis = as.factor(data$diagnosis)
data$diagnosis = revalue(data$diagnosis, c("control" = "0", "schizophrenia" = "1"))

#Creating model with diagnosed predicted by pitch range 
model1 = glmer(diagnosis ~ 1 + range + (1+trial|participant) + (1|study), family= "binomial", data = data)
summary(model1)

#Confusion matrix
data$PredictionsPerc=predict(model1)
data$Predictions[data$PredictionsPerc>0]="0"
data$Predictions[data$PredictionsPerc<=0]="1"
confusionMatrix(data = data$Predictions, reference = data$diagnosis, positive = "1") 

#Making predictions a factor so we can calculate their sensitivity, specificity, PPV and NPV in the loop - alternative to extracting values out of confusionmatrix 
data$Predictions = as.factor(data$Predictions)
sensitivity(data = data$Predictions, reference = data$diagnosis, positive = "1") 
specificity(data = data$Predictions, reference = data$diagnosis, negative = "0") 
posPredValue(data = data$Predictions, reference = data$diagnosis, positive = "1") 
negPredValue(data = data$Predictions, reference = data$diagnosis, negative = "0") 

#ROC curve - area under the curve 
rocCurve <- roc(response = data$diagnosis, predictor = data$PredictionsPerc)
auc(rocCurve) 
ci (rocCurve)
plot(rocCurve, legacy.axes = TRUE) 

```


Loop action
```{r}

#Participant as factor and creating folds_id variable to give participants numbers from 1 to 61
data$participant = as.factor(data$participant)
data$folds_id = as.numeric(data$participant)

#Cross-validation
folds = createFolds(unique(data$participant), k = 4)

#Creating empty lists to put data into
accuracy = NULL
specificity = NULL
sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL
n = 1

#Creating loop
for(i in folds){
  #subsetting testdata
  testsubset = subset(data, folds_id %in% i)
  
  #subsetting traindata
  trainsubset = subset(data,!(folds_id %in% i))
  
  #training the model on the traindata subset
  trainmodel = glmer(diagnosis ~ 1 + range + (1+trial|participant) + (1|study), family= "binomial", data = trainsubset)
  
  #Confusionmatrix
  testsubset$PredictionsPerc = predict(trainmodel, testsubset, allow.new.levels = T)
  testsubset$Predictions[testsubset$PredictionsPerc>0]="0"
  testsubset$Predictions[testsubset$PredictionsPerc<=0]="1"
  cm = confusionMatrix(data = testsubset$Predictions, reference = testsubset$diagnosis, positive = "1")
  testsubset$Predictions = as.factor(testsubset$Predictions)
  
  #Extracting values from confusion matrix
  accuracy[n] = cm$overall["Accuracy"]
  specificity[n] = cm$byClass["Specificity"]
  sensitivity[n] = cm$byClass["Sensitivity"]
  PPV[n] = cm$byClass["Pos Pred Value"]
  NPV[n] = cm$byClass["Neg Pred Value"]
  
  #ROC curve - area under the curve 
  roccurve = roc(response = testsubset$diagnosis, predictor = testsubset$PredictionsPerc)
  AUC[n] = auc(roccurve) 
  ci(roccurve)
  plot(roccurve, legacy.axes = TRUE) 
  
  n=n+1
}

df = data.frame(accuracy, sensitivity, specificity, PPV, NPV, AUC)

```



### Question 2 - Which single acoustic predictor is the best predictor of diagnosis?

```{r}

#Creating empty lists to put data into
accuracy = NULL
specificity = NULL
sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL
n = 1

#Creating loop
for(i in folds){
  #subsetting testdata
  testsubset = subset(data, folds_id %in% i)
  
  #subsetting traindata
  trainsubset = subset(data,!(folds_id %in% i))
  
  #training the model on the traindata subset
  trainmodel = glmer(diagnosis ~ 1 + rqa_L + (1+trial|participant) + (1|study), family= "binomial", data = trainsubset)
  
  #Confusionmatrix
  testsubset$PredictionsPerc = predict(trainmodel, testsubset, allow.new.levels = T)
  testsubset$Predictions[testsubset$PredictionsPerc>0]="0"
  testsubset$Predictions[testsubset$PredictionsPerc<=0]="1"
  cm = confusionMatrix(data = testsubset$Predictions, reference = testsubset$diagnosis, positive = "1")
  testsubset$Predictions = as.factor(testsubset$Predictions)
  
  #Extracting values from confusion matrix
  accuracy[n] = cm$overall["Accuracy"]
  specificity[n] = cm$byClass["Specificity"]
  sensitivity[n] = cm$byClass["Sensitivity"]
  PPV[n] = cm$byClass["Pos Pred Value"]
  NPV[n] = cm$byClass["Neg Pred Value"]
  
  #ROC curve - area under the curve 
  roccurve = roc(response = testsubset$diagnosis, predictor = testsubset$PredictionsPerc)
  AUC[n] = auc(roccurve) 
  ci(roccurve)
  plot(roccurve, legacy.axes = TRUE) 
  
  n=n+1
}

df1 = data.frame(accuracy, sensitivity, specificity, PPV, NPV, AUC)

```


### Question 3 - Which combination of acoustic predictors is best for diagnosing schizophrenia?

Now it's time to go wild! Use all (voice-related) variables and interactions you can think of. Compare models and select the best performing model you can find.

Remember:
- Out-of-sample error crucial to build the best model!
- After choosing the model, send Celine and Riccardo the code of your model

```{r}

#Creating empty lists to put data into
accuracy = NULL
specificity = NULL
sensitivity = NULL
PPV = NULL
NPV = NULL
AUC = NULL
n = 1

#Creating loop
for(i in folds){
  #subsetting testdata
  testsubset = subset(data, folds_id %in% i)
  
  #subsetting traindata
  trainsubset = subset(data,!(folds_id %in% i))
  
  #training the model on the traindata subset
  trainmodel = glmer(diagnosis ~ 1 + range + radius + (1+trial|participant) + (1|study), family= "binomial", data = trainsubset)
  
  #Confusionmatrix
  testsubset$PredictionsPerc = predict(trainmodel, testsubset, allow.new.levels = T)
  testsubset$Predictions[testsubset$PredictionsPerc>0]="0"
  testsubset$Predictions[testsubset$PredictionsPerc<=0]="1"
  cm = confusionMatrix(data = testsubset$Predictions, reference = testsubset$diagnosis, positive = "1")
  testsubset$Predictions = as.factor(testsubset$Predictions)
  
  #Extracting values from confusion matrix
  accuracy[n] = cm$overall["Accuracy"]
  specificity[n] = cm$byClass["Specificity"]
  sensitivity[n] = cm$byClass["Sensitivity"]
  PPV[n] = cm$byClass["Pos Pred Value"]
  NPV[n] = cm$byClass["Neg Pred Value"]
  
  #ROC curve - area under the curve 
  roccurve = roc(response = testsubset$diagnosis, predictor = testsubset$PredictionsPerc)
  AUC[n] = auc(roccurve) 
  ci(roccurve)
  plot(roccurve, legacy.axes = TRUE) 
  
  n=n+1
}

df2 = data.frame(accuracy, sensitivity, specificity, PPV, NPV, AUC)

```


### Question 4: Properly report the results

METHODS SECTION: how did you analyse the data? That is, how did you extract the data, designed the models and compared their performance?

RESULTS SECTION: can you diagnose schizophrenia based on voice? which features are used? Comment on the difference between the different performance measures.

### Bonus question 5

You have some additional bonus data involving speech rate, pauses, etc. Include them in your analysis. Do they improve classification?

### Bonus question 6

Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them.
